Locally-Oriented Programming: A stunningly simple programming model for GPU programming in Fortran and C++

### 2011-11-02 papers/LOPe/intro.tex

In this paper we examine a language-based paradigm that allows the compiler to
take on a larger portion of the PCP burden.  Anyone programming in OpenCL or
CUDA is aware that explicit loop structures over array elements in a serial
program are removed and replaced by a kernel program that is run over all of
the

>>>>

In this paper we examine a language-based paradigm that allows the compiler to
take on a larger portion of the PCP burden.  Anyone programming in OpenCL or
CUDA is aware that explicit loop structures over array elements in a serial
program are removed and replaced by a kernel program that is run over all of the
elements of the input arrays.  We propose Locally Orientated Programming
extensions (LOPe) to the Fortran and C languages that formally adopt this
programming model.

### 2011-11-02 papers/LOPe/motivation.tex

>>>>

The LOPe programming model restricts the programmer --- while implementing a stencil
algorithm --- to a local view of the array index space.  Within a concurrent function,
only the local array element plus a small halo region is visible to the programmer.
This reduces complexity by all boundary conditions and processor topology information
from the algorithm implementation.  This reduction in complexity reduces programming
errors.  While developing the convolution example described later in the paper we
made an indexing error in applying the 2D stencil loops in the standard serial 
Fortran test implementation.  This error required over 3 hours of programming time
to repair.  First the error had to be isolated to the function implementing the
convolution (it was first thought to be in the complicated tiff image output routine
as the convolution code was ``thought'' to be too simple to wrong.  Then the index
error has to be understood.  As will be seen, LOPe makes it more difficult to make
these errors as Fortran array intrics can be used.  In some instance, the restricted
semantics of LOPe allows the compiler to catch errors (e.g., some errors involving
race conditions).
